---
title: "Analysis"
description: |
  A new article created using the Distill format.
author:
  - name: Jonah Fisher
    url: https://github.com/JonahF95
    affiliation: Harvard T.H. Chan Department of Biostatistics
    affiliation_url: https://www.hsph.harvard.edu/biostatistics/
date: "`r Sys.Date()`"
output: 
  distill::distill_article:
    toc: true
    toc_depth: 6
    toc_float: true
    css: toc-styles.css
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(data.table)
library(ggsci)
library(ewastools)
library(caret)
library(bnstruct)

theme_set(theme_bw())
```

# Load in qc data

```{r load data}
dat <- fread("pdqc.csv")
meth <- readRDS("pp/meth.rds")
```

# Create matrix of predictors and get beta matrix

M values are a log transformation of the beta methylation values. Log transformed
gives them a better distribution. Instead of ranging from 0-1 as in beta values
the 0.5 value becomes a 0 for m values and higher/lower levels of methylation become
positive/negative respectively.

```{r matrix of predictors}
# Predictor matrix to perform PCA on
xmat <- do.call("rbind", list(meth$ctrlG, meth$ctrlN, meth$ctrlR)) %>% t
colnames(xmat) <- c(paste0("ctrlg", 1:850), paste0("ctrln", 1:850), paste0("ctrlr", 1:850))
missmat <- which(is.na(xmat), arr.ind = T)
xmat[missmat] <- rowMeans(xmat, na.rm = T)[missmat[, 1]]
xpc <- prcomp(xmat)

#M values for outcome
m <- meth %>% dont_normalize() %>% wateRmelon::Beta2M()
centm <- m %>% t %>% scale(scale = F) %>% t
```

# Linear modeling

The simplest form to start is with linear modeling. Principal components 1-10 of
the predictor matrix is used to predict the mean centered DNA methylation
expression

```{r predict values}
set.seed(260)
irow <- sample(65:nrow(centm), size = 1000)
lmout <- list()
for(i in irow){
  lmout[[paste0("row", i)]] <- lm(y ~ ., data = data.table(y = centm[i,], x = xpc$x[, 1:10]))
}
```

## Pull out different coefficients and significance

We look to see if there's any consistency between the coefficients and the statistical
significance of each of the 10 PCA predictors

### Coefficients

```{r get coefficients}
coefdt <- map(lmout, ~coef(.x)[2:11]) %>% do.call("rbind", .) %>% data.table
colnames(coefdt) <- str_extract(colnames(coefdt), "PC\\d*")
coefdt <- melt(coefdt)

ggplot(coefdt, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_hline(linetype = "dashed", yintercept = 0) +
  labs(x = "Predictor", y = "Coefficient distribution")
ggsave("coefficient_predictors.png")

#Only PC1/PC2 seems to have coefficients significantly different from 0 consistently
t.test(coefdt[variable == "PC1", value])
t.test(coefdt[variable == "PC2", value])

coefdt[, t.test(value), by = variable]
```

### Significance

```{r get significance}
sigdt <- map(lmout, ~summary(.x)$coefficients[-1, 3]) %>% do.call("rbind", .) %>% data.table
colnames(sigdt) <- str_extract(colnames(sigdt), "PC\\d*")
sigdt <- melt(sigdt)

ggplot(sigdt, aes(x = variable, y = value)) +
  geom_boxplot() +
  geom_hline(lty = "dashed", yintercept = 0) +
  labs(x = "Predictor", y = "Test statistic distribution")
ggsave("tstat.png")

# t.test(sigdt[variable == "PC1", value])
# t.test(sigdt[variable == "PC2", value])
sigdt[, t.test(value), by = variable]
```

### R squared

There are fairly high r2 values from these models

```{r get r2}
r2 <- map(lmout, ~summary(.x)$r.squared) %>% data.table

ggplot(r2, aes(x = unlist(.))) +
  geom_histogram(bins = 20, color = "black", fill = "grey") +
  labs(x = "Multiple R2 values")
ggsave("rsq.png")

```


## RMSE

```{r get rmse for the models}
map(lmout, ~sd(fitted(.x))) %>% unlist -> sds
map(lmout, ~sqrt(mean(.$residuals^2))) %>% unlist -> rmses

formse <- data.table(sds = sds, rmses = rmses)
formse[, mean(sds)]
formse[, mean(rmses)]

ggplot(formse, aes(x = sds, y = rmses)) +
  geom_point() +
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4)) +
  geom_smooth(method = "lm") +
  labs(x = "SDs of fitted values", y = "RMSE of model")
ggsave("sdvsrmse.png")
```


