---
title: "Get Methylation Data"
description: |
  A new article created using the Distill format.
author:
  - name: Jonah Fisher
    url: https://github.com/JonahF95
    affiliation: Harvard T.H. Chan Department of Biostatistics
    affiliation_url: https://www.hsph.harvard.edu/biostatistics/
date: "`r Sys.Date()`"
output: distill::distill_article
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(data.table)
library(rvest)
```

# Get data

## Read in html page for sample list and pull replicate pairs

```{r get data together}
url <- "https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE61496"
l <- read_html(url)
l <- l %>% html_text %>% str_split("\n") %>% unlist
idx <- str_detect(l, "_[A|B]") %>% which

dat <- data.table(
  id = l[idx - 1],
  num = l[idx]
)

dat[, c("num", "orig") := tstrsplit(num, "_")]
setkey(dat, num)
```

## Get urls for raw files

There's a similar FTP address for the files that we can plug into a looping function
which takes the GSM identifier of the file and downloads the corresponding idat

```{r get idats}
if(!dir.exists("idats")){
  dir.create("idats")
}

#Where this series is located
repourl <- "https://ftp.ncbi.nlm.nih.gov/geo/samples/GSM1506nnn/"

for(i in dat[, id]){
  #Need to grab more info on name of sample to download
  ftplines <- paste0(repourl, i, "/suppl/") %>% readLines() 
  
  #Get red and green idat file locations
  idatnames <- ftplines[str_detect(ftplines, "\\.idat\\.gz")] %>% tstrsplit('\\"', keep = 2) %>% unlist
    paste0(repourl, i, "/suppl/", idatnames[[1]]) %>% 
    download.file(destfile = paste0("idats/",idatnames[[1]]))
    
    paste0(repourl, i, "/suppl/", idatnames[[2]]) %>% 
    download.file(destfile = paste0("idats/",idatnames[[2]]))
}
```

## Pull information on biological sex for qc check

All samples for this study have have bio sex listed on the website after a uniform pattern
of "sex, 1=m, 2=f: [number]" which we can use to match to web scraped results.  

```{r}
s <- readLines("https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSM1506327")
biosex <- map(dat[, id], function(acc){
  readLines(paste0("https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=", acc)) %>%
  str_extract("sex, 1=m, 2=f: \\d") %>%
  .[!is.na(.)] %>% 
  str_split("") %>% 
  unlist() %>% 
  tail(n=1)
})

dat[, sex := c("m", "f")[unlist(biosex) %>% as.numeric]]

fwrite(dat, "samplesheet.csv")
```













